---
title: "ggplot2_basics"
date: "17/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1: First steps with ggplot

## Setup

### Load ggplot2


1. Load `ggplot2` using the `library` function

```{r}
```


### Read data

2. Read the data file called "data_ex1_1.csv" in the "data" folder and store it in an object (name it as you like).

```{r}

```


### Explore data

3. Look at the data structure using `str` function. How many columns are there ? What are their type ?
```{r}



```

4. Compute the mean for each modality of the `group` column. 

```{r}
```

## Visualize the data using ggplot2

We want to have a look at the distribution of the data (column "value").

5. Run the following lines to create a ggplot object, store it in an object called `my_plot`.

```{r}
my_plot <- ggplot(data = data_ex1_1, aes(x = 0, y = value)) # We don't need to put " as ggplot2 recpgnizes the columns' names

```

Take a look at the `my_plot` object by running `my_plot`.

6. Add a layer `geom_boxplot` to the `my_plot` object. Name this new plot `my_boxplot`, and take a look at it. Do you see anything particular ?

```{r}

```



7. Add a layer `geom_point` to the `my_boxplot` object.

Do you see anything ? 
```{r}

my_boxplot+
  geom_point() 
```

Except if you're a witcher, you should not be able do distinguish anything :-).



8. Let's guide you to make things clearer. 
In the `geom_point` function, add these arguments:

* `alpha = 0.3`
* `position = position_jitter(width = .1, height = 0)`

What is the role of each argument ?

```{r}
```

Anyway, things are clearer, but one aspect of the dataset may pass under the radar.

9. Suppose we would like to have a boxplot for each modality of the `group` column. What should we specify instead of `x = 0` in the `aes` ? Redo the steps 5 to 8, with one boxplot by group. What can you conclude about the distribution of `value` within the groups ? 


```{r}

```

10. Instead of `geom_boxplot`, use the `geom_violin` function

```{r}


```

What can you conclude about the importance of visualization ?



# Exercise 2: other visualizations

1. Load the `msleep` dataset (`data("msleep")`), take a look at `?msleep` to know more about the columns

```{r}
```


2. Look at a histogram of the sleep_total column. Add a vertical line (see `geom_vline` at the mean of the `sleep_total` column) Roughly speaking, how long do the mammals sleep ? Do the values vary a lot ? 

```{r}

```

3. Use a `geom_bar` to look at the number of mammals of each `vore` type. How many mammals have their `vore` type not specified ?

```{r}


```

4. Use `geom_smooth()` and `geom_point()` layers to look at the relationship between `sleep_total` and `awake` columns. What do you see ? How can you explain this wonderful relationship :o

```{r}
```



......




Congratulations, you've won a Nobel prize. When an animal does not sleep, he is awake :-).


5.A bit more serious now, any relation ship between the body weight and the brain weight ?


```{r}
```

6. What make things difficult to see ? Which transformation could you apply to look at the relationship between these two variables ? Apply that transformation (*Hint:* no need to create new columns; simply put the function around the `bodywt` and `brainwt` inside the `aes` argument).

```{r}


```

7. Add horizontal and vertical lines (`geom_hline` and `geom_vline`), to make things easier to read (lines should cross at the (0,0) point)

```{r}


```

8. Could you roughly derive a (mathematical) relationship between the body weight and the brain weight ? 


9. Confirm your thoughts using the `lm` function

```{r}


```

