---
title: "ggplot2_basics"
author: "Rémi Mahmoud"
date: "17/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

### Load ggplot2


1. Load `ggplot2` using the `library` function

```{r}
library(ggplot2)
```


### Read data

2. Read the data file called "data_ex1_1.csv" in the "data" folder and store it in an object (name it as you like).

```{r}


# data_ex1_1 <- tibble(value =  c(rnorm(200, mean = 2, sd = sqrt(0.5)), rnorm(100, 0,sqrt(0.5)), rnorm(100,4,sqrt(0.5))), group = rep(c("A", "B"), each = 200))

# readr::write_csv(data_ex1_1, "data/data_ex1_1.csv")
data_ex1_1 <- read.csv("data/data_ex1_1.csv")
```


### Explore data

3. Look at the data structure using `str` function. How many columns are there ? What are their type ?
```{r}

str(data_ex1_1)


```

4. Compute the mean for each modality of the `group` column. 

```{r}

mean(data_ex1_1[which(data_ex1_1$group == "A"),1] )
mean(data_ex1_1[which(data_ex1_1$group == "B"),1] )
```

## Visualize the data using ggplot2

We want to have a look at the distribution of the data (column "value").

5. Run the following lines to create a ggplot object, store it in an object called `my_plot`.



```{r}
my_plot <- ggplot(data = data_ex1_1, aes(x = 0, y = value)) # We don't need to put " as ggplot2 recpgnizes the columns' names

```

Take a look at the `my_plot` object by running `my_plot`.

6. Add a layer `geom_boxplot` to the `my_plot` object. Name this new plot `my_boxplot`, and take a look at it. Do you see anything particular ?

```{r}

my_boxplot <- my_plot +
  geom_boxplot()

my_boxplot

```



7. Add a layer `geom_point` to the `my_boxplot` object.

Do you see anything ? 
```{r}

my_boxplot+
  geom_point() 
```

Except if you're a witcher, you should not be able do distinguish anything :-).

8. Let's guide you to make things clearer. 
In the `geom_point` function, add these arguments:

* `alpha = 0.3`
* `position = position_jitter(width = .1, height = 0)`

What is the role of each argument ?

```{r}

my_boxplot+
  geom_point(alpha = .3, position = position_jitter(width =.1, height = 0)) 
```

Anyway, things are clearer, but one aspect of the dataset may pass under the radar.

9. Suppose we would like to have a boxplot for each modality of the `group` column. What should we specify instead of `x = 0` in the `aes` ? Redo the steps 5 to 8, with one boxplot by group. What can you conclude about the distribution of `value` within the groups ? 


```{r}

my_plot2 <- ggplot(data = data_ex1_1, aes(x =group, y= value)) +
  geom_boxplot()+
  geom_point(alpha = .3, position = position_jitter(width =.1, height = 0)) 
my_plot2

```

10. Instead of `geom_boxplot`, use the `geom_violin` function

```{r}

my_plot2 <- ggplot(data = data_ex1_1, aes(x =group, y= value)) +
  geom_violin()+
  geom_point(alpha = .3, position = position_jitter(width =.1, height = 0)) 
my_plot2

```

What can you conclude about the importance of visualization ?



# Exercise 2: other visualizations

1. Load the `msleep` dataset (`data("msleep")`), take a look at `?msleep` to know more about the columns

```{r}
data("msleep")
?msleep
```


2. Look at a histogram of the sleep_total column. Add a vertical line (see `geom_vline` at the mean of the `sleep_total` column) Roughly speaking, how long do the mammals sleep ? Do the values vary a lot ? 

```{r}
ggplot(data = msleep, aes(x = sleep_total)) +
  geom_histogram(alpha =.4, fill = "blue") +
  geom_vline(xintercept = mean(msleep$sleep_total)) 

ggplot(data = msleep, aes(x = sleep_total)) +
  geom_density(alpha =.4, fill = "blue") +
  geom_vline(xintercept = mean(msleep$sleep_total)) 

```

3. Use a `geom_bar` to look at the number of mammals of each `vore` type 

```{r}

ggplot(data=  msleep, aes(x = vore)) +
  geom_bar()

```

4. Use `geom_smooth` and `geom_point()` layers to look at the relationship between `sleep_total` and `awake` columns. What do you see ? How can you explain this wonderful relationship :o

```{r}
ggplot(data =msleep, aes(x = sleep_total, awake)) +
  geom_point() +
  geom_smooth(method = "lm")
```



......




Congratulations, you've won a Nobel prize. When an animal does not sleep, he is awake :-).


5.A bit more serious now, any relation ship between the body weight and the brain weight ?


```{r}
msleep %>% 
ggplot(data =., aes(x = brainwt, bodywt)) +
  geom_point() +
  geom_smooth() 
```

6. What make things difficult to see ? Which transformation could you apply to look at the relationship between these two variables ? Apply it (no need to create new columns; simply put the function around the `bodywt` and `brainwt` inside the `aes` argument).

```{r}


msleep %>% 
ggplot(data =., aes(x = log(brainwt), log(bodywt))) +
  geom_point() +
  geom_smooth(method = "lm") 

```

7. Add horizontal and vertical lines (`geom_hline` and `geom_vline`), to make things easier to read (lines should cross at the (0,0) point)

```{r}


msleep %>% 
ggplot(data =., aes(x = log(brainwt), log(bodywt))) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0)

```

8. Could you roughly derive a (mathematical) relationship between the body weight and the brain weight ? 

$\log(y) = 5 + 1.3 \times \log(x)$
$y = e^5 \times x^1.3$ 


9. Confirm your thoughts using the `lm` function

```{r}

lm(log(bodywt) ~log(brainwt), data =msleep)

```

