---
title: "ggplot2_customization"
date: "17/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1: Basic customization

1. Load the `iris` dataset and plot (scatterplot) the Petal.Width against the Petal.Length of the flowers. Add a linear smoothing. Each species should have its own linear smoothing .

```{r}
ggplot(data = iris, aes(x = Petal.Width, y=  Petal.Length, color = Species)) +
  geom_point() +
  geom_smooth(method = "lm")
```

2. Add the `theme_bw` (bw stands for black & white) to your plot.

```{r}
ggplot(data = iris, aes(x = Petal.Width, y=  Petal.Length, color = Species)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_bw() 

```

3. Add the `theme_dark` after the `theme_bw`. What happens ? 

```{r}
ggplot(data = iris, aes(x = Petal.Width, y=  Petal.Length)) +
  geom_point() +
  theme_bw() +
  theme_dark()

```



4. Many themes exist. List them by running `ls("package:ggplot2")[grepl(ls("package:ggplot2"),pattern = "theme")]`.

```{r}
ls("package:ggplot2")[grepl(ls("package:ggplot2"),pattern = "theme")]

```

5. Instead of specifying the theme in each plot, you can set the active theme by using the `theme_set` function. Set `theme_bw` as the default theme for each plot.

```{r}

theme_set(theme_bw())

```

6. Specify the colors of each species using the `scale_colour_manual` function. Pick colors of your choice (all R colors are available using the `colours()` function). Look at the help page of the function to get the syntax. 

```{r}
ggplot(data = iris, aes(x = Petal.Width, y=  Petal.Length, color = Species)) +
  geom_point() +
  scale_color_manual(values = c("sienna", "purple", "red"))


```

7. Add a title to your plot. Put it in bold

```{r}
ggplot(data = iris, aes(x = Petal.Width, y=  Petal.Length, color = Species)) +
  geom_point() +
  scale_color_manual(values = c("sienna", "purple", "red")) +
  theme(plot.title = element_text(face = "bold")) +
  labs("Nice title")

```


8. Put the axis titles in italic


```{r}
ggplot(data = iris, aes(x = Petal.Width, y=  Petal.Length, color = Species)) +
  geom_point() +
  scale_color_manual(values = c("sienna", "purple", "red")) +
  theme(plot.title = element_text(face = "bold"), axis.title = element_text(face = "italic")) +
  labs(title = "Nice title")

```

9. Change the color and the type of the background grid (`panel.grid`, which is modified using `element_line`)

10. Put the legend at the bototm of the plot.
```{r}
p <- ggplot(data = iris, aes(x = Petal.Width, y=  Petal.Length, color = Species)) +
  geom_point() +
  scale_color_manual(values = c("sienna", "purple", "red")) +
  theme(plot.title = element_text(face = "bold"),
        axis.title = element_text(face = "italic"),
        panel.grid = element_line(color = "grey", linetype = "dashed"),
        legend.position = "bottom") +
  labs(title = "Nice title")

p
```

11. Play with other parameters of the plot. Change the margin of the legend (`legend.margin` + `margin` function).

```{r}


p +
  theme(plot.background = element_rect(fill = "lightblue", linetype = 4), legend.margin = margin(2,2,2,2, "mm"))
```

# Exercise 2: other things to look at

1. Load the `mtcars` dataset.

```{r}
data("mtcars")

```


2. Plot values of `wt` versus `mpg`. Add a `geom_text_repel` (or `geom_label_repel`) layer (from the package `ggrepel`), that will show the name of each car on each point. Each number of cylinders should have a different color (labels and points)

```{r}

p_cars <- ggplot(data = mtcars, aes(x = mpg, y= wt, color = factor(cyl))) +
  geom_point() +
  theme_bw() +
  ggrepel::geom_text_repel(aes(label = rownames(mtcars)))

p_cars
```


3. Add a `facet_wrap` layer to separate these plots by cylinder.

4. Change the size of the axis' ticks (see `axis.ticks.length` argument of `theme` combined with the `unit` function). Play with the `strip.background` argument of `theme`. What does it change ? 

```{r}

p_cars <- p_cars +
  facet_wrap(cyl~.) +
  theme(axis.ticks.length = unit(0.5,"cm"), strip.background = element_rect(fill = "azure"))
```


5. Do another plot (of your choice) with the `mtcars` dataset. Combine the two plots using the `plot_grid` function (package `cowplot`).

```{r}

pcars_2 <- ggplot(mtcars, aes(x = factor(gear), y = qsec)) +
                    geom_violin(alpha = 0.1) +
  geom_point(position = position_jitter(width = .05, height = 0)) +
  theme_bw() +
  labs(x = "Gear")

pcars_2

two_plots_cars <- cowplot::plot_grid(p_cars, pcars_2)
```

6. Save the plot using the `ggsave` function. Explore the arguments

```{r, eval= FALSE}


# ggsave("exercise/two_plots.png", height = 10, width = 12, unit = "cm")
```

